title: "Coursera: Survival Analysis with R"
author: "Yujin Zhang"
date: "7/6/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

library(survival)
library(survminer)
library(ggplot2)
```

##Running a KM plot and log-rank test
First, we load the dataset and do some variable manipulation.

```{r}
#load course data
tbl1 = read_csv(file = "./data/prcatice_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = as.factor(gender)) %>% 
  mutate(ethnicgroup = as.factor(ethnicgroup)) %>% 
  mutate(copd = as.factor(copd))

skimr::skim(tbl1) #explore the dataset

```

Then we run overall KM method.
```{r}
km_fit = survfit(Surv(fu_time, death) ~ 1,data = tbl1) #call overall k-m plot
plot(km_fit)

summary(km_fit, times = c(1:7,30,60,90*(1:10)))
```

Run KM method by gender
```{r}
km_gender = survfit(Surv(fu_time, death) ~ gender,data = tbl1) #split the curve by gender
plot(km_gender)

survdiff(Surv(fu_time, death) ~ gender, data = tbl1) #compare survival curve

```

##Running Simple Cox Model

```{r}
cox_mod = coxph(Surv(fu_time, death) ~ ethnicgroup, data = tbl1)
summary(cox_mod)
```

About missing data:  
1. should be missing at random (example of survey data/not random)  
2. the process to treat missing data: understand why those data are missing- try to fill in the missing values on the basis of another variable- complete case analysis/ mean substitution/ multiple imputation/ maximum likelihood  

##Running Multiple Cox Model
```{r}
cox_mod2 = coxph(Surv(fu_time, death) ~ ethnicgroup+age+gender+copd+prior_dnas, data = tbl1)
summary(cox_mod2)

```

What to do when model does not converge:
```{r}
tbl2 =
  tbl1 %>% 
  mutate(quintile = as.factor(quintile))

cox_uncon = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_uncon)

# 1. change the reference category
tbl2 =
  tbl2 %>% 
  mutate(quintile = relevel(quintile, ref = 2)) #quintile 1 as the ref cat 

cox_uncon = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_uncon) 
#still not converge

# 2. combine categories
tbl2 = 
  tbl1 %>% 
  mutate(quintile_5grp = quintile) %>% 
  mutate(quintile_5grp = replace(quintile_5grp, quintile_5grp == "0", "5"),
         quintile_5grp = factor(quintile_5grp))

cox_con = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_con) 

# 3. drop the quintile zero(set to null)
tbl2 = 
  tbl1 %>% 
  mutate(quintile_5grp = quintile) %>% 
  mutate(quintile_5grp = replace(quintile_5grp, quintile_5grp == "0", NA),
         quintile_5grp = factor(quintile_5grp))

cox_con = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_con) 
```

##Checking the proportionality assumption

```{r}
#default arguments: cox.zph(fit, transform = "km", global = TRUE)

fit = coxph(Surv(fu_time, death) ~ gender, data = tbl2)
temp = cox.zph(fit)

print(temp)
plot(temp)
 
```

The line is pretty flat, meaning that the effect of gender little during the follow-up. That's good news.  
Also, we can use 'ggcoxzh()', which produces graphs pf the scaled Schoenfield residuals against the transformed time for each covariate.

```{r}
#deviance residuals

res.cox = coxph(Surv(fu_time, death) ~ age, data = tbl2)
ggcoxdiagnostics(res.cox, type = "deviance",
                 linear.predictions = FALSE, ggtheme = theme_bw())

#test linear relation
ggcoxfunctional(Surv(fu_time, death) ~ age + log(age) + sqrt(age), data = tbl2)
```

##Proportionality assumption not met

If there is a statistical interaction between gender and time for example. Then the model is short of a coefficient.
```{r}
#test statistically significant
fit = coxph(Surv(fu_time, death) ~ gender + tt(gender), data = tbl2) # 'tt' -> time-transform function
summary(fit)

```

##Choose predictors

Apply backwards method


