title: "Coursera: Survival Analysis with R"
author: "Yujin Zhang"
date: "7/6/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(ggplot2)
```

##Running a KM plot and log-rank test
First, we load the dataset and do some variable manipulation.

```{r}
#load course data
tbl1 = read_csv(file = "./data/prcatice_data.csv") %>% 
  mutate(gender = as.factor(gender)) %>% 
  mutate(ethnicgroup = as.factor(ethnicgroup)) %>% 
  mutate(copd = as.factor(copd))

head(tbl1)
```

Then we run overall KM method.
```{r}
km_fit = survfit(Surv(fu_time, death) ~ 1,data = tbl1) #call overall k-m plot
plot(km_fit)

summary(km_fit, times = c(1:7,30,60,90*(1:10)))
```

Run KM method by gender
```{r}
km_gender = survfit(Surv(fu_time, death) ~ gender,data = tbl1) #split the curve by gender
plot(km_gender)

survdiff(Surv(fu_time, death) ~ gender, data = tbl1) #compare survival curve

```

##Running Simple Cox Model

```{r}
cox_mod = coxph(Surv(fu_time, death) ~ ethnicgroup, data = tbl1)
summary(cox_mod)
```

About missing data:  
1. should be missing at random (example of survey data/not random)  
2. the process to treat missing data: understand why those data are missing- try to fill in the missing values on the basis of another variable- complete case analysis/ mean substitution/ multiple imputation/ maximum likelihood  

##Running Multiple Cox Model
```{r}
cox_mod2 = coxph(Surv(fu_time, death) ~ ethnicgroup+age+gender+copd+prior_dnas, data = tbl1)
summary(cox_mod2)

```

What to do when model does not converge:
```{r}
tbl2 =
  tbl1 %>% 
  mutate(quintile = as.factor(quintile))

cox_uncon = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_uncon)

# 1. change the reference category
tbl2 =
  tbl2 %>% 
  mutate(quintile = relevel(quintile, ref = 2)) #quintile 1 as the ref cat 

cox_uncon = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_uncon) 
#still not converge

# 2. combine categories
tbl2 = 
  tbl1 %>% 
  mutate(quintile_5grp = quintile) %>% 
  mutate(quintile_5grp = replace(quintile_5grp, quintile_5grp == "0", "5"),
         quintile_5grp = factor(quintile_5grp))

cox_con = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_con) 

# 3. drop the quintile zero(set to null)
tbl2 = 
  tbl1 %>% 
  mutate(quintile_5grp = quintile) %>% 
  mutate(quintile_5grp = replace(quintile_5grp, quintile_5grp == "0", NA),
         quintile_5grp = factor(quintile_5grp))

cox_con = coxph(Surv(fu_time, death) ~ age + gender + copd + quintile + ethnicgroup, data = tbl2)
summary(cox_con) 
```

